\documentclass[12pt,letterpaper]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Force pdflatex processing even with "$ latex" (required by arXiv)
%\pdfoutput=1 %compatiblity with LaTeX online
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{xcolor}
\definecolor{nicered}{rgb}{0.7,0.1,0.1}
\definecolor{nicegreen}{rgb}{0.1,0.5,0.1}
\usepackage[colorlinks=true,citecolor= nicegreen,linkcolor= nicered]{hyperref}
\usepackage{graphicx}
\usepackage{cancel}
\usepackage{datetime}

\setlength{\textwidth}{180mm}
\setlength{\oddsidemargin}{-1cm}
\setlength{\evensidemargin}{-1cm}
\setlength{\textheight}{22cm} 
\setlength{\topmargin}{-1cm}
\title{Standard Model}

\author{Enrico Nardi\\
(Notes by students of the course)\thanks{\currenttime; \today.}}
\date{}

\begin{document}
\maketitle

\section{Natural units}
We use $\hbar=c=1$.  
Space-time is based in the Group $SO(1,3)$, with the conversion factor between the ``3'' and ``1'' part being $c$. 

\section{Review of Standard Model}
Based on symmetry $SU(3)\times SU(2)\times U(1)$, with 8 gluons, 3 $W$  bosons and one Abelian gauge boson. When the symmetry is broken we have $W_{\mu}; Z_{\mu}$. The spin--$1/2$ particle content is
\begin{align}
  l=
  \begin{pmatrix}
    \nu_L\\ e_{L}
  \end{pmatrix}\sim& (1,2,-1/2)  & e\sim&(1,1,-1)\\
  Q=
  \begin{pmatrix}
    u_L\\ d_L
  \end{pmatrix}\sim& (3,2,1/6)  & u\sim&(3,1,2/3)\\
&&d\sim&(3,1,-1/3)\\
  H=
  \begin{pmatrix}
    \phi^+\\ \phi^0
  \end{pmatrix}\sim& (1,2,1/2)  & &
\end{align}
\begin{itemize}
\item The 8 gluons are in the adjoint of $SU(3)$
\item The 3 $W$ are in the adjoint of $SU(2)$
\item $B_{\mu}$ is a singlet.
\end{itemize}
The Lagrangian is
\begin{align*}
  \mathcal{L}=&-\frac{1}{4}G_{\mu\nu}^{\alpha}-\frac{1}{4}W^{a}_{\mu\nu}W_{a}^{\mu\nu}
-\frac{1}{4}B_{\mu\nu}B^{\mu\nu}\\
&-\frac{g_3^2}{64\pi^2}\theta_3\epsilon_{\mu\nu\rho\sigma}G^{\alpha\mu\nu}G^{\alpha\rho\sigma}\\
&+\text{terms which can be eliminate similar to the previous one}
\end{align*}
where
\begin{align*}
  G_{\mu\nu}^{\alpha}=\partial_{\mu}...
\end{align*}
Moreover
\begin{align*}
  \theta_3=\theta_3+2\pi\,.
\end{align*}

\begin{align*}
  \mathcal{L}_{gf}=&\bar{l}_i \gamma^{\mu}\mathcal{D}_{\mu}l_i
+\bar{e}_i \gamma^{\mu}\mathcal{D}_{\mu}e_i
+\bar{Q}_i \gamma^{\mu}\mathcal{D}_{\mu}Q_i
+\bar{u}_i \gamma^{\mu}\mathcal{D}_{\mu}u_i
+\bar{d}_i \gamma^{\mu}\mathcal{D}_{\mu}d_i
+\bar{N}_i \gamma^{\mu}\partial_{\mu}N_i
\end{align*}
\begin{align*}
  \mathcal{L}_{H}=\left( \mathcal{D}_{\mu}H \right)^{\dagger}
\left( \mathcal{D}^{}{\mu}H \right)-V(H^{\dagger}H)
\end{align*}
\begin{align*}
  \mathcal{L}=-\bar{l}_i y^l_{ij}e_j H
-\bar{Q}_i y^u_{ij}u_j H
-\bar{Q}_i y^d_{ij}d_j \tilde{H}
-\bar{N}_i y^N_{ij}N_j \tilde{H}
\end{align*}
The parameters are
\begin{itemize}
\item Gauge sector: $g_1,g_2,g_3,\theta_3$ (4)
\item Yukawa (ME): 3+3+3+3+1=13
\item Higgs: $\mu^2;\lambda$ (2)
\end{itemize}

\section{Problems of Standard Model}

\begin{align*}
  [\mathcal{L}]=&M^4
\end{align*}
The action
\begin{align*}
  S=&\int d^4x \mathcal{L} & [S]=&0
\end{align*}
In $D$-dimensions

\begin{align*}
  [\mathcal{L^{(D)}}]=&M^D
\end{align*}
The Action
\begin{align*}
  S=&\int d^Dx \mathcal{L}^{(D)} & [S]=&0
\end{align*}
\begin{align*}
  \left[ \partial_{\mu}\phi \partial^{}{\mu}\phi\right]=&M^D\\
\left[ \bar{\psi} \gamma^{\mu}\psi\right]=& M^D\\
\left[ F^{\mu\nu} F_{\mu\nu}\right]...
\end{align*}
Then
\begin{align*}
  [\phi^n(\bar{\psi }\psi)^m]=&M^{n+3m}\\
\end{align*}
Otras dimensiones
\begin{itemize}
\item Gauge couplings $[g_j]=0$
\item Yukawa couplings: $[y]=0$
\item Higgs potential parameters: $[\mu]=M^1;\qquad [\lambda]=0$
\end{itemize}
A renormalizable theory
\begin{align*}
  4\ge [\text{dimension of parameters}]\ge 0\,,
\end{align*}
The existence of infinities is required because locality.
\begin{align*}
  \int d^4x\mathcal{L}_{\phi}\sim \int d^4 x\phi^{3}
\end{align*}
Ultaviolet divergences are expected due to locality. In string theory the amplitudes are finite. 

In SM, it is possible to calculate
\begin{align*}
  \mathcal{L}^{(D-5)}_{\mu}\sim&\frac{F_2(0)}{2m_e}\bar{e}\sigma_{\mu\nu}e F^{\mu\nu}\\
&\bar{\mu}(p')\left[ i\frac{\sigma^{\mu\nu}}{2m} \right]\bar{\mu}(p)
\end{align*}


If the fundamental theory is known, the effective Lagrangian terms can be explained in terms of a more fundamental theory. 

But seen to very low energy with only the four light quarks, the Lagrangian does not includes the muon decay. The decay width for $\tau$ decay
\begin{align*}
  \bar{u}_{\nu_{\tau}}(l)\left[ \frac{ig}{\sqrt{2}}\gamma^{\mu}\frac{1}{2}(1-\gamma_5) \right]u_{\tau}(k)\frac{-i}{r^2-M_W^2+i\epsilon}\left[ g_{\mu\nu}+
\frac{(\xi-1)r_{\mu}-r_{\nu}}{r^2-\xi M_W^2} \right]
\bar{u}_j(p)\left[ \frac{ig}{\sqrt{2}}\gamma^{\mu}\frac{1}{2}(1-\gamma_5) \right]v_i(q)
\end{align*}
The effective width is
\begin{align*}
  A_{\tau^{-}\text{decay}}=-i
\frac{g^2}{8M_W^{2}}V_{ij}
\left[\bar{u}_{\nu_{\tau}}\gamma^{\mu}(1-\gamma_5)u_{\tau}  \right]
\left[\bar{u}_j\gamma^{\mu}(1-\gamma_5) v_i  \right]
\end{align*}
So that
\begin{align*}
  \mathcal{L}^{\text{eff}}=-i\frac{G_F}{\sqrt{2}}V_{ij}
\left[\bar{u}_{\nu_{\tau}}\gamma^{\mu}(1-\gamma_5)u_{\tau}  \right]
\left[\bar{u}_j\gamma^{\mu}(1-\gamma_5) v_i  \right]
\end{align*}
is a dimension-six Lagrangian. 

In other situation, where we are now with the SM, were the fundamental
theory is not known. However, we can add to the SM Lagrangian,
effective terms which the same gauge invariance.
\begin{align*}
  \mathcal{L}=\mathcal{L}_{\text{ME}}+\mathcal{L}^{\text{eff}}
\end{align*}
where
\begin{align*}
  \mathcal{L}^{\text{eff}}=\sum_I C_I \theta^{4+n}(\text{light fields})
\end{align*}
where
\begin{align*}
  c_I=\frac{\tilde{c}_I}{\Lambda^n}
\end{align*}
the $\Lambda$-scale is discouraging large.

In Donogue book, a QED calculation is done for $E\ll m_e$ with path integral
\begin{align*}
  Z=\int \left[ D\Phi \right]\left[ D\phi \right]
e^{i\int d^{x}\left[\mathcal{L}(\phi,\phi)+\mathcal{L}(\phi,\Phi)+\mathcal{L}(\Phi,\Phi)  \right]}
\end{align*}
where $\Phi$ are heavy fields, we arrive to
\begin{align*}
  \mathcal{L}^{\text{eff}}=\frac{\alpha^{2}}{90m_e^4}\left[ \left( F_{\mu\nu}F^{\mu\nu} \right)^2
+\frac{7}{16}\left( F_{\mu\nu}\widetilde{F}^{\mu\nu} \right)^2 \right]
\end{align*}
What is the scale of new physics?, without additional parameters the scale of new physics associated with
\begin{align*}
  \frac{1}{\Lambda}(LH)^2\,,
\end{align*}
is $\Lambda\sim 10^{14}$, while the others
\begin{align*}
  \frac{1}{\Lambda^2}\bar{\psi}\psi\bar{\psi}\psi
\end{align*}
the scale is around $100$~TeV

\section{Dimension-0}
Let
\begin{align*}
  \mathcal{L}_0=\frac{1}{2}\partial_{\mu}\varphi\partial^{\mu}\varphi-\frac{1}{2}m^{2}\varphi\varphi-C_{\varphi}
\end{align*}
Therefore
\begin{align*}
  E_0=C_{\varphi}\int_V d^3x+\frac{1}{2}\int
  \frac{d^3\mathbf{p}}{E_{\mathbf{p}}(2\pi)^3}
E_{\mathbf{p}}(2\pi)^3\delta(0)
\end{align*}
Normal ordering:
\begin{align*}
  \left[ a_{\mathbf{p}},a_{\mathbf{p}'}^{*} \right]=2 E_{\mathbf{p}}
(2\pi)^3\delta(\mathbf{p}-\mathbf{p}')
\end{align*}
Divergences appear at $x\to \infty$ and $\mathbf{p}\to\infty$, and
\begin{align*}
  \Omega=\int_V d^3x e^{i(\mathbf{p}=0)}=(2\pi)^3\delta(0)
\end{align*}
with
\begin{align*}
  \rho=E_0/\Omega
\end{align*}
we have
\begin{align*}
  \rho=C_{\varphi}+\frac{1}{16\pi}\left\{
    \underbrace{\Lambda^4+m_{\varphi}\Lambda^2}_{\text{UV}}+
    \underbrace{\frac{1}{4}m^4\log \frac{m^{2}}{\Lambda^2}
+\frac{1}{4}m^4 \left( \frac{1}{2}\log 4 \right)}_{\text{IR
      problem $\propto m_{\varphi}^{2}$}}+m^4\mathcal{O}\left( m^2/\Lambda^2 \right) \right\}
\end{align*}
Including the fermion contribution
\begin{align*}
  E_0^{\psi}=C_{\psi}\int_V d^3x-\sum_{\sigma}\frac{1}{2}\int
  \frac{d^3\mathbf{p}}{E_{\mathbf{p}}(2\pi)^3}
E_{\mathbf{p}}(2\pi)^3\delta(0)
\end{align*}
and
\begin{align*}
  \rho=C_{\psi}-\frac{1}{8\pi}\left\{
    \underbrace{\Lambda^4+m_{\psi}\Lambda^2}_{\text{UV}}+
    \underbrace{\frac{1}{4}m_\psi^4\log \frac{m_\psi^{2}}{\Lambda^2}
+\frac{1}{4}m_\psi^4 \left( \frac{1}{2}\log 4 \right)}_{\text{IR
      problem $\propto m_{\psi}^{2}$}}+m_\psi^4\mathcal{O}\left( m_\psi^2/\Lambda^2 \right) \right\}
\end{align*}

For the full Lagrangian
\begin{align*}
  \rho=C_{\varphi}+C_{\psi}-\frac{1}{16\pi}\left\{
    [m_{\varphi}-m_{\psi}]\Lambda^2+
\cdots  \right\}
\end{align*}
For the top mass (or a SUSY scale around 100~GeV)
\begin{align*}
  \frac{\delta\rho_t}{\rho_{\text{exp}}}\sim 
  \left( \frac{1.7\times 10^{11}\ \text{eV}}{2.3\times 10^{-3}\
      \text{eV}} \right)^4
       \sim 10^{55}
\end{align*}

\section{Dimension-2}

\textbf{Hierarchy Problem}, related with the hierarchy of the Higgs
mass.
\begin{align*}
  V(|\phi|^2)=&\lambda \left( |\phi|^2-v^2 \right)\nonumber\\
            =&\lambda|\phi|^4-\mu^2|\phi|^2
\end{align*}
The contribution from $W,Z,\gamma$ to the Higgs-mass at one-loop is
\begin{align*}
  \frac{1}{16\pi^2}g^2\Lambda^2
\end{align*}
the autointeraction contribution is
\begin{align*}
  \frac{1}{16\pi^2}\lambda^2\Lambda^2
\end{align*}
The top contribution is
\begin{align*}
    -\frac{1}{16\pi^2}\lambda_t^2\Lambda^2
\end{align*}


If $\Lambda_{\text{New Physics}}=10\ $TeV and $m_H=200\ \text{GeV}$
\begin{align}
  \frac{1}{16\pi^2}g^2\Lambda^2\sim &(700\text{GeV})^2\\
  \frac{1}{16\pi^2}\lambda^2\Lambda^2\sim & (500\text{GeV})^2\\
-\frac{1}{16\pi^2}\lambda_t^2\Lambda^2\sim & -(2000\text{GeV})^2\\
\end{align}

and
\begin{align*}
  m_H^2=\left( m_H^0 \right)^2-[-100+10+5]\left( 200\ \text{GeV} \right)^2
\end{align*}

The precise result in terms of the parameter $\mu$ is
\begin{align*}
  \delta \mu^2 =\frac{3}{16\pi^2v^2}
\left[2m_W^2+m_z+m_H-4 m_t^2  \right]\Lambda^2
\end{align*}
the cancellation of the terms inside brackets happen if
\begin{align*}
  m_H^2\sim (307\ \text{GeV})^2
\end{align*}
which is known as the Veltman's condition. 

With RN neutrinos
\begin{align*}
  M_{\nu}=
  \begin{pmatrix}
    0 & \lambda v\\    
    \lambda v &M \\    
  \end{pmatrix}
\end{align*}
with two eigenvalues
\begin{align*}
  m_{\nu}\approx \lambda^2\frac{v^2}{M}\nonumber\\
 M_N\approx M-\lambda^2\frac{v^2}{M}
\end{align*}
for $m\sim 0.1\ $eV, and $\lambda\sim 0.1$
\begin{align*}
  M\sim 10^{12}\ \text{GeV}
\end{align*}
Since RH neutrinos have couplings with the Higgs
\begin{align*}
  -\frac{1}{2}\lambda \overline{N}L\Phi
\end{align*}
The diagram $\Phi$--$\Phi$ at one loop gives the contribution
\begin{align*}
  -i\int
  \frac{d^p}{(2\pi)^4}\frac{i}{\cancel{p}}\frac{i}{\cancel{p}-m}
=&i(i)\int
\frac{\pi^2}{16\pi^4}\frac{1}{2}|p|^2d|p|^2\frac{\cancel{p}_E}{-p_E^2}
\frac{\cancel{p}_E+M}{-p_E^2-M^2}\nonumber\\
=&\frac{-i}{16\pi^2}\int_{M_H}^{\Lambda^{2}}(1-M^2/p^2)dp^{2}\nonumber\\
=&-i \frac{1}{16\pi^2}\left[ \lambda^2-M^2\log \frac{\Lambda^2}{m_H^2} \right]
\end{align*}
As $\Lambda>N$, the Higgs mass acquires a 1-loop contribution of order $M$. 

\section{Electroweack symmetry breaking models}

\subsection{Only Higgs}
\subsection{Supersymmetry}

\subsection{Technicolor}


\subsection{Extradimensions}
\begin{enumerate}
\item Large compact extradimensions
\item Warped extra dimensions 
\end{enumerate}
Let $R$ the compactification radius. Let two particles of mass $M$ and
$m$ two particles separated some distance $R$

If $r\ll R$ 
\begin{align*}
\nabla \cdot \frac{\mathbf{F}}{m}=&G_D \rho\nonumber\\
\frac{{F}}{m}r^{n+2}=& G_D M
\end{align*}
for $r\gg R$ 
\begin{align*}
\frac{F}{m}R^n=&eG_D M \nonumber\\
\frac{F}{m}r^2=&\frac{G_D}{R^n} M 
\end{align*}
\begin{align}
  G_{\text{Newton}}=&1/M_p^2 \nonumber\\
G_D=&1/M_D^{2+n}
\end{align}
\begin{align*}
G_N=\frac{G_D}{R^n}\rightarrow R=\left( \frac{M_p^2}{M_D^{2+n}} \right)^{1/n}
=\left( \frac{M_p^2}{M_D^2} \right)^{1/n}M_D^{-1}.
\end{align*}
and the dimensional could be at mm-level. 

How test in colliders?

\begin{align}
  \hat{g}_{\hat{\mu}\hat{\nu}}=\eta_{\hat{\mu}\hat{\nu}}+h_{\hat{\mu}\hat{\nu}}
\end{align}
where 
\begin{align}
h_{\hat{\mu}\hat{\nu}}=
\begin{pmatrix}
  h_{\mu\nu} & A_{\mu i}\\
A_{\nu i} &\Phi_{ij}
\end{pmatrix}
\end{align}
\begin{align*}
   h_{\mu\nu}(x,y)=\sum_{\mathbf{h}}h_{\mu\nu}^{\mathbf{n}}e^{\frac{i2\pi}{R}\mathbf{n}\cdot \mathbf{y}}\qquad y\in [0,R]
\end{align*}

In Warped extradimensions.

The metric is
\begin{align*}
  ds^2=\underbrace{dy^2}_{\text{5-dim}}
+ e^{-2 k|y|}(dr^2+r^2d\Omega^2)
\end{align*}
Similar to the FRW metric. 

Two branes at $y-0$ and $y=r_c\pi$

The action is
\begin{align*}
  S_{\text{Grav}}=\int d^4x\int dy\sqrt{-G}\left\{ 
-\Lambda+2^3M \right\}
\end{align*}
After some calculations
\begin{align*}
  \left\{ \cdots \right\}
=\left\{ \bar{g}^{\mu\nu}(D_{\mu})^{\dagger}(D_{\nu}H)
-\lambda \left[ \left| H \right|^2 -e^{-2k r_c\pi}v_0^2\right] \right\}
\end{align*}
If $v_0\sim M_p$, is enough o supression factor with $2\pi r_c \pi\sim
50$, a proper $v$ at electroweak scale is obtained.

\section{Dimension-four}

\subsection{Triviality/Instability}
\begin{align*}
  \frac{1}{\alpha(\Lambda^2)}\to 0\qquad\text{si $\alpha(\Lambda^2)\to\infty$}
\end{align*}
if the couplings goes infinite before the scale of validity of the
theory, then the only option is that the corresponding coupling be
zero at the low energy scale.

None of the three gauge  couplings of the standard model have problems
with triviality.
\begin{align*}
  \mu^2 \frac{\partial h^u}{\partial \mu^2}\approx &
\frac{1}{16\pi^2}h^u_{33}\left[ ... \right]
\end{align*}
with a negative beta function and there is not problems with $\Lambda$-pole.

For $\lambda$
\begin{align*}
  \mu^2 \frac{d\lambda}{d\mu^2}=
\frac{1}{16\pi^{2}}\cdot 3\cdot 3
\left[ 4\lambda^2+4\lambda|h^{\dagger}|^2-(|h^{\dagger}|)^2 \right]
\end{align*}
If the case of Landau pole $\Lambda_{LP}<\Lambda_P\rightarrow m_H>130\text{GeV}$

With a Higgs of a mass of $125\text{GeV}$, then the $\Lambda_{LP}>\Lambda_{\text{GUT}}$.

\subsection{Maxwell equations}
Because we are close to the stability model but for the while
everything is consistent.

\begin{align}
  \label{eq:hom_m_eq}
  \boldsymbol{\nabla}\cdot\mathbf{B}&=0,&\boldsymbol{\nabla}\times \mathbf{E}+\frac{\partial\mathbf{B}}{\partial t}&=0\\
  \label{eq:inhom_m_eq}
  \boldsymbol{\nabla}\cdot\mathbf{E}&=\rho,&\boldsymbol{\nabla}\times \mathbf{B}-\frac{\partial\mathbf{E}}{\partial t}&=\mathbf{J}\,.
\end{align}
which is gauge invariant. the Maxwell equations can be written  terms of the vector and scalar potentials.

\subsubsection{Aarnov-Bohm  effect}
The Hamiltonian for a system in presence of a magnetic field
\begin{align*}
  \mathcal{H}=\frac{1}{2m}\left( \mathbf{p-e\mathbf{A}} \right)+
e\phi+V
\end{align*}
and
\begin{align*}
  \mathcal{H}\psi=i \frac{\partial \psi}{\partial t}
\end{align*}
with solution
\begin{align*}
  \psi=e^{i e} \int_0^r\mathbf{A}'d\mathbf{r}' \psi'. 
\end{align*}
The experimental result is
\begin{align*}
  \delta\Phi=\delta\Phi^0+e\oint \mathbf{A}\cdot d\mathbf{r}
\end{align*}

In the non-Abelian case it is possible to make a \emph{pure gauge} transformation such that
\begin{align*}
  A_{\mu}'\to U^{-1}\partial _{\mu}U
\end{align*}
If we take a trivial vacuum and apply to it some gauge transformation
\begin{align*}
  |1\rangle=U|0\rangle
\end{align*}
Therefore
\begin{align*}
    |n+1\rangle=U|n\rangle
\end{align*}

To define the vacuum in a proper way
\begin{align*}
|0\rangle_\theta=  \sum_{h=-\infty}^{\infty}e^{i\theta_n}|n\rangle
\end{align*}

However the QCD Lagrangian must allow a term
\begin{align*}
  \theta \frac{g^2}{32\pi^2}\int d^4 x F_{\mu\nu}^a\tilde{F}^{a\mu\nu}
\end{align*}
\subsection{Yukawas}
\begin{align*}
  \mathcal{L}=\bar{\psi}_{Q}(i\cancel{D})\psi+\cdots
\end{align*}
where
\begin{align*}
  \psi_Q=
\begin{pmatrix}
Q_1\\
Q_2\\
Q_3
\end{pmatrix}
\end{align*}
Therefore
\begin{align*}
  \psi\to U\psi
\bar{\psi}\to \bar{\psi} U^\dagger
\end{align*}
where $U$ is a $U(3)_Q$ symmetry. The same for the other terms. The full symmetry of the Yukawa Lagrangian os
\begin{align*}
  U_Q(3)\times U_u(3)\times U_d(3)
\end{align*}
or
\begin{align*}
  \left[ U(1) \right]^3\times SU(3)_Q\times SU(3)_u\times SU(3)_d
\end{align*}
The Yukawa Lagrangian is
\begin{align*}
  \mathcal{L}=\bar{Q}_i Y_Q^{ij} u_j+\cdots
\end{align*}
In some basis where $y_u$ will be diagonal with one complex entry, and
hierarchical entries. One part of the puzzle is why the other matrix
$Y_d$ is a approximately diagonal.

\section{Higher dimension operators}
\subsection{Weinberg operator}
\begin{align}
  \frac{f_{ab}}{\Lambda}\left( \overline{l^c_{ia}}l_{jb} \right)\left( \Phi_k\Phi_l \right)\epsilon_{ik}\epsilon_{jl}
\end{align}
\subsubsection{Tree level}
\begin{enumerate}
\item $\psi\Phi\rightarrow$ fermion singlet of $SU(2)$

  \begin{align}
    &\frac{1}{\Lambda}\left( \phi^0\nu-\phi^+e \right)\left( \phi^0\nu-\phi^+e \right) \nonumber\\
    \sim& \frac{1}{\Lambda}\phi^0\nu \phi^0\nu\sim \frac{v^2}{\Lambda}\nu\nu 
  \end{align}
From the tree level diagram with coupling $\lambda$
\begin{align}
  m_{\nu}=&\lambda \lambda \frac{v^2}{M_N}& \Lambda\sim \frac{1}{M_N}
\end{align}

\item $\bar{\psi}\psi\rightarrow$ scalar triplet of $SU(2)$
  \begin{align}
    \frac{1}{\Lambda}\left(\underbrace{\phi^0\phi^0}_{-1/2-1/2}\nu\nu
    -\underbrace{\phi^+\phi^0}_{1/2-1/2}\nu\nu
    +\underbrace{\phi^+\phi^+}_{1/2+1/2}\nu\nu
  \right)
  \end{align}
From the tree level diagram with couplings $\lambda$ and $\mu$
\begin{align}
  m_{\nu}=&\mu \lambda \frac{v^2}{M_\Delta}& \Lambda\sim \frac{\mu}{M_\Delta}
\end{align}

\item $\psi\Phi\rightarrow$ fermion triplet 
From the tree level diagram with couplings $\lambda$ and $\mu$
\begin{align}
  m_{\nu}=&\mu \lambda \frac{v^2}{M_\Sigma}& \Lambda\sim \frac{1}{M_\Sigma}
\end{align}

\end{enumerate}
\subsubsection{1 loop level}
Sthephano Bertollini et al have studied unification of SM in $SO(10)$,
which contains the right handed neutrinos in the $16$ (15 fermions and
one right-handed neutrino), \footnote{$E_6$ contains 27 vector-like
  fermions} and one intermedaite scale $\Lambda$ compatible with
type-I seesaw.

$SU(5)$ contains triplet real
%%%%%%%%%%%%%%%%%
\bibliographystyle{h-physrev4}%apsrev4-1long
\bibliography{susy}
\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "smdp"
%%% End: 
